Selective stock enhancement of an exploited species undermines community stability

### 2021-12-06 rmd/manuscript.Rmd
edit: manuscript

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to sustainable management of natural resources. Captive breeding is a common practice to conserve or enhance wild populations of diverse plant and animal taxa. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles), this method is still pervasive in conservation and fisheries. In fisheries for example, XXX have been spent annually to release billions of hatchery-reared individuals globally. This is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks. The current debate, however, overlooks the fact that we have rarely assessed the community-wide impact of selective stock enhancement. Species are all embedded in the complex web of interacting organisms, and excessive release of particular species may have unforeseen consequences for ecological communities in the long term. Evidence for this hypothesis is lacking, however.

@@

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities. Increasing numbers of released individuals $R$ led to greater temporal fluctuations of communities as illustrated by increased CV (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals (increased reproducing individuals). Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with increased numbers of released individuals (XX). The SDs showed a similar trend, but the relationship was flatter than the mean at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios.

Importantly, the reduced mean of community density may arise as long as humans control the number of releases. The stability of ecological communities rests on internal density regulation that allows for stable coexistence of competing species. For example, significant niche difference in resource use ($\alpha_{ij, i \ne j} < 1$) enables competing species to grow from low population densities because the numerically dominant species undergoes stronger intraspecific competition. However, the number of release is not subject to density-dependent feedback. The excessive individuals released by humans impose competition that would otherwise not exist and undermine the self-regulation process of stable communities. This argument is supported by the fact that the destabilizing effect of selective stock enhancement manifested even with weak interspecific competition that allows for stable coexistence of competing species. Perhaps, the only exception would be a scenario in which the species of interest is on the verge of local extinction with little or no density regulation (i.e., very large carrying capacity).

## Empirical evidence

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon on long-term stability of stream fish communities. Masu salmon represents one of the most valuable freshwater resources in Hokkaido Island, Japan, and XX leads extensive hatchery programs across the island. In protected watersheds where any form of exploitation is prohibited (including recreational angling), long-term monitoring has been conducted with consistent sampling methods over the past two decades along with official records of the number of fish released. Therefore, the data set the stage for a "natural experiment" to test the theoretical predictions. In this study, we assembled time-series data of stream fish communities at XX sites within XX protected watersheds (see Methods for selection criterion). Using hierarchical Bayesian models, we quantified the relationship between community dynamics and the number of fish released while accounting for other potential confounding factors.

@@

Our study system satisfies key theoretical assumptions that are required for the destabilizing effect of selective stock enhancement to emerge. The first assumption is that released fish compete with wild fish. This type of competition is very likely to occur in the study streams. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that this species compete with other species of stream fish communities. In particular, masu salmon raised in hatchery environments are more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild. Second, our theory predicts the emergence of destabilizing effect when the number of fish released is externally controlled. This is exactly what was observed in our study system. In the protected watersheds, there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). Instead, these released fish are likely controlled by socioeconomic factors, such as budget allocation to hatchery programs in this region. As such, released fish are probably "excessive" given the density-dependent regulation of natural communities, imposing intense resource competition that otherwise would not exist.

@@

Although captive breeding programs carefully consider the fitness of captive-bred individuals in the wild, the debate often ignores the capacity of recipient ecosystems to support biodiversity. The results of this study suggest that the ignorance of the ecosystem capacity is critical, undermining the long-term stability of natural communities. Our research has two important implications to improve captive breeding programs. First, release of captive-bred individuals should be avoided when strong density-dependence operates in the recipient ecosystem. Second, we should prioritize the preservation and restoration of recipient ecosystems, which are rapidly degraded due to myriads of human influences. Without this integrative management framework that recognizes the fundamental process of biodiversity maintenance, selective stock enhancement will never be effective but will impair biodiversity.

>>>>

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to sustainable management of natural resources. Captive breeding is a common practice to conserve or enhance wild populations of diverse plant and animal taxa. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles), this method is still pervasive in conservation and fisheries. In fisheries for example, XXX have been spent to produce and release XXX billions of hatchery-reared individuals globally. The widespread use of captive breeding is perhaps built upon the "myth" that numerical benefits of released individuals may exceed the associated risks. The current debate, however, overlooks the fact that we have rarely assessed the community-wide impact of selective stock enhancement. Species are all embedded in the complex web of interacting organisms, and excessive release of particular species may have unforeseen consequences for ecological communities in the long term. Evidence for this hypothesis is lacking, however.

@@

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities. Increasing numbers of released individuals $R$ led to greater temporal fluctuations of communities as illustrated by increased CV (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals (increased reproducing individuals). Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter than the mean at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios.

Importantly, the reduced mean of community density may arise as long as humans control stock enhancement. The stability of ecological communities rests on internal density regulation that allows for stable coexistence of competing species. For example, significant niche difference in resource use ($\alpha_{ij, i \ne j} < 1$) enables competing species to grow from low population densities because the numerically dominant species undergoes stronger intraspecific competition. However, the number of release is not subject to density-dependent feedback. The excessive individuals released by humans impose additional competition that undermines the self-regulation process of stable communities. This argument is supported by the fact that the destabilizing effect of selective stock enhancement manifested even with weak interspecific competition that otherwise allows for stable coexistence of competing species. Perhaps, the only exception would be a scenario in which the species of interest is on the verge of local extinction with little or no density regulation, i.e., very large carrying capacity relative to the current population density.

To demonstrate the relevance of our general theory to natural systems, we assessed the potential impacts of selective stock enhancement of masu salmon on long-term stability of stream fish communities. Masu salmon represents one of the most valuable freshwater resources in Hokkaido Island, Japan, and XX leads extensive hatchery programs across the island. In protected watersheds where any form of exploitation is prohibited (including recreational angling), long-term monitoring has been conducted with consistent sampling methods over the past two decades along with official records of stock enhancement. Therefore, the data set the stage for a "natural experiment" to test the theoretical predictions. In this study, we assembled time-series data of stream fish communities at XX sites within XX protected watersheds (see Methods for selection criterion). Using hierarchical Bayesian models, we quantified the relationship between community dynamics and the number of fish released while accounting for effects of climates and local abiotic factors.

@@

Our study system satisfies key theoretical assumptions that are required for the destabilizing effect of selective stock enhancement to emerge. The first assumption is that released fish compete with wild fish. This type of competition is very likely to occur in the study streams. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that this species compete with other species of stream fish. In particular, masu salmon raised in hatchery environments are more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild. Second, our theory predicts the emergence of destabilizing effect when the number of fish released is externally controlled. This is exactly what was observed in our study system. In the protected watersheds, there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). Instead, these released fish are likely controlled by socioeconomic factors, such as budget allocation to hatchery programs in this region. As such, released fish are probably "excessive" given the density-dependent regulation of natural communities, imposing intense resource competition that otherwise would not exist.

@@

Although captive breeding programs carefully consider the fitness of captive-bred individuals in the wild, the debate ignores the capacity of recipient ecosystems to support biodiversity. The results of this study suggest that the ignorance of the ecosystem capacity is critical, undermining the long-term stability of natural communities. Our research has two important implications to improve captive breeding programs. First, release of captive-bred individuals should be avoided when strong density-dependence operates in the recipient ecosystem. XXX Second, we should prioritize the preservation and restoration of recipient ecosystems, which are rapidly degraded due to myriads of human influences. XXX Without this integrative framework that recognizes the fundamental process of biodiversity maintenance, selective stock enhancement will never be effective but impairs biodiversity.
